---
title: "Chi-Maps"
author: "Nicolas Beglinger, Nicolas Schmidheiny"
date: "12/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Initialize
```{r, echo = FALSE, message = FALSE}
pkgTest <- function(x) 
  {
    if (!require(x, character.only = TRUE))
    {
      install.packages(x, dep = TRUE)
        if(!require(x, character.only = TRUE)) stop("Package not found")
    }
}

pkgTest("sf")
pkgTest("tidyverse")
pkgTest("tmap")

#for palette explorer
pkgTest("tmaptools")
pkgTest("shiny")
pkgTest("shinyjs")


crs_lv95  <- "+init=epsg:2056"

tmap_mode("view")
source("functions.r")
```

```{r}
accidents <- st_read("shapefiles/accidents/verkehrsunfaelle_stadt_zh_points_new.shp")
accNoHighway <- filter(accidents, RdTyp_n != "Motorway" & RdTyp_n != "Expressway")
```

Subsamples of accidents, e.g. Bicycle, Pedestrians, Severe etc.
```{r}
fatalities = filter(accidents, AccdSC_ == "Accident with fatalities")
severe = filter(accidents, AccdSC_ == "Accident with severe injuries")
light = filter(accidents, AccdSC_ == "Accident with light injuries")
propDam = filter(accidents, AccdSC_ == "Accident with property damage")
severeFatal = filter(accidents, AccdSC_ == "Accident with fatalities" | AccdSC_ == "Accident with severe injuries")

bicycle = filter(accidents, AccdnIB == "true")
pedestrians = filter(accidents, AccdnIP == "true")
motorCycle = filter(accidents, AccdnIM == "true")

bicycleSevere = filter(bicycle, AccdSC_ == "Accident with severe injuries")
pedestriansSevere = filter(pedestrians, AccdSC_ == "Accident with severe injuries")
motorCycleSevere = filter(motorCycle, AccdSC_ == "Accident with severe injuries")

```

## Creating Hexbinning Count Maps of Accidents(All), Bicycle, Motorcycle and Pedestrians
```{r}

AllCountMap <- countMap(MyHexBin, data = accidents, cellsize = 100, title = "All Accidents Counts Map", datapoints = accidents)

BicCountMap <- countMap(MyHexBin, data = bicycle, cellsize = 100, title = "Bicycle Accidents Counts Map", datapoints = bicycle)

MotCountMap <- countMap(MyHexBin, data = motorCycle, cellsize = 100, title = "Motorcycle Accidents Counts Map", datapoints = motorCycle)

PedCountMap <- countMap(MyHexBin, data = pedestrians, cellsize = 100, title = "Pedestrians Accidents Counts Map", datapoints = pedestrians)

```

## Visualize Hexbinning Count Maps of Accidents(All), Bicycle, Motorcycle and Pedestrians
```{r}

AllCountMap
BicCountMap
MotCountMap
PedCountMap

```

## Creating Hexbinning Chi Maps of Bicycle, Motorcycle and Pedestrians
```{r}

BicChiMap <- chiMap(MyHexBin, reference = accNoHighway, focus = bicycle, cellsize = 100, title = "Bicycle Accidents Chi Map", accidentspoints = NULL, mode = "view") 

MotChiMap <- chiMap(MyHexBin, reference = accidents, focus = motorCycle, cellsize = 100, title = "Motorcycle Accidents Chi Map", accidentspoints = NULL, mode = "view") 

PedChiMap <- chiMap(MyHexBin, reference = accNoHighway, focus = pedestrians, cellsize = 100, title = "Pedestrians Accidents Chi Map", accidentspoints = NULL, mode = "view")

BicSevChiMap <- chiMap(MyHexBin, reference = bicycle, focus = bicycleSevere, cellsize = 200, title = "Bicycle Severe Accidents Chi Map", accidentspoints = accidents, mode = "view")

```

## Visualizing Hexbinning Chi Maps of Bicycle, Motorcycle and Pedestrians 
```{r}

BicChiMap
MotChiMap
PedChiMap
BicSevChiMap

```

## Creating Hexbinning Chi Maps of Severity-Categories
```{r}

SevereChiMap <- chiMap(MyHexBin, reference = accidents, focus = severe, cellsize = 100, title = "Severe Accidents Chi Map", datapoints = severe, mode = "view") 

```

## Visualizing Hexbinning Chi Maps of Severity-Categories
```{r}

SevereChiMap

```

## Histogram 
```{r}

BicCount <- MyHexBin(bicycle, extent = st_bbox(accNoHighway))
AllCount <- MyHexBin(accNoHighway, extent = st_bbox(accNoHighway))

BicAllRatio <- BicCount %>% mutate(Bic = BicCount$ïAccUID, All = AllCount$ïAccUID, Ratio = Bic / All) %>%  filter(!is.na(Ratio)) %>%  filter(Ratio > 0.2) %>%  filter(Bic >= 10)

hist(BicAllRatio$Ratio, breaks = 29)
hist(BicAllRatio$Bic, breaks = 29)

tm_shape(BicAllRatio) +
  tm_polygons(
        col = "Ratio",
        id = "Bic",
        palette = c("#F7F7F7", "#F6B596", "#CB4A42", "#67001F"),
        alpha = 0.7,
        style = "pretty",
        lwd = 0.1
      ) +
tm_shape(accNoHighway) +
  tm_dots(
        size = 0.01,
        col = 'black'
        ) +
tm_shape(bicycle) +
  tm_dots(
        size = 0.01,
        col = 'red'
        )


```
