---
title: "Chi-Maps"
author: "Nicolas Beglinger"
date: "12/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, message = FALSE}
pkgTest <- function(x) 
  {
    if (!require(x, character.only = TRUE))
    {
      install.packages(x, dep = TRUE)
        if(!require(x, character.only = TRUE)) stop("Package not found")
    }
}

pkgTest("sf")
pkgTest("tidyverse")
pkgTest("tmap")

#for palette explorer
pkgTest("tmaptools")
pkgTest("shiny")
pkgTest("shinyjs")


crs_lv95  <- "+init=epsg:2056"

tmap_mode("view")
```

```{r}
accidents <- st_read("shapefiles/accidents/accidentsOnRoads_ZH.shp")
```

Subsamples of accidents, e.g. Bicycle, Pedestrians, Severe etc.
```{r}
fatalities = filter(accidents, AccdSC_ == "Accident with fatalities")
severe = filter(accidents, AccdSC_ == "Accident with severe injuries")
light = filter(accidents, AccdSC_ == "Accident with light injuries")
propDam = filter(accidents, AccdSC_ == "Accident with property damage")
severeFatal = filter(accidents, AccdSC_ == "Accident with fatalities" | AccdSC_ == "Accident with severe injuries")

bycicle = filter(accidents, AccdnIB == "true")
pedestrians = filter(accidents, AccdnIP == "true")
motorCycle = filter(accidents, AccdnIM == "true")
```

## Chi Maps

MyHexBin function - First create the hexagons
```{r}
MyHexBin <- function(x, extent = NA, cellsize = 100, mrg = 100)
{
    
    if (is.na(extent)) {
      require(sf)
      x1 <- st_coordinates(x)[,1]
      y2 <- st_coordinates(x)[,2]
      minx1 <- min(x1) - mrg
      maxx1 <- max(x1) + mrg
      miny2 <- min(y2) - mrg
      maxy2 <- max(y2) + mrg
    } else {
      minx1 <- extent[1] - mrg
      maxx1 <- extent[3] + mrg
      miny2 <- extent[2] - mrg
      maxy2 <- extent[4] + mrg
    }

    # coords defines the bounding rectangle of the hexagonal tessellation,
    # including the margin defined by parameter mrg.
    # Alternatively, coords might be defined by the coordinates of a particular 
    # window that one might be interested in. Would require changing this code.
    coords <- list(rbind(c(minx1,miny2), c(minx1,maxy2),
                         c(maxx1,maxy2), c(maxx1,miny2), c(minx1,miny2)))
    # Turn bounding coords into an sfc_POLYGON object and inherit CRS from x
    pg = st_sfc(st_polygon(coords), crs = st_crs(x))

    # Create a hexagonal grid --> sfc_POLYGON object
    hexPol <- st_make_grid(pg, cellsize = cellsize, square = FALSE)

    hexPol_npts <- aggregate(x[1], hexPol, length) %>%
                  replace(is.na(.), 0) #%>% rename(npts_hex = id)

    return(hexPol_npts)
}
```

countMap function
```{r}

countMap = function(hexBinFunction, data, cellsize = 100, title, datapoints, pointopacity = 1, mode = "view") {
    
  hexData = MyHexBin(data, st_bbox(accidents), cellsize = cellsize, mrg = 100) %>%
  filter(!ïAccUID == 0)
  
  quantDiff = quantile(hexData$ïAccUID, probs = seq(0,1,0.0025))
  
  if (mode == "view") {
    tm_shape(datapoints) +
      tm_dots(
        size = 0.01,
        alpha = pointopacity
      ) +
    tm_shape(hexData) +
      tm_polygons(
        col = "ïAccUID",
        id = "ïAccUID",
        palette = c("#F7F7F7", "#F6B596", "#CB4A42", "#67001F"),
        alpha = 0.7,
        style = "fixed",
        breaks = c(min(hexData$ïAccUID),
               quantDiff[320], 
               quantDiff[360],
               quantDiff[376],
               quantDiff[393],
               quantDiff[397],
               quantDiff[400],
               max(hexData$ïAccUID)),
        lwd = 0.1
      ) +
    tm_layout(title = title)
    
  } else if (mode == "return") {
      return(hexData)
    
  } else {
      warning("Mode has to be \"view\" or \"return\"!")
  }
}

```

chiMap function
```{r}
chiMap = function(hexBinFunction, reference, focus, cellsize, title, datapoints, pointopacity = 1, mode = "view") {
  hexRef = hexBinFunction(reference, st_bbox(reference), cellsize = cellsize, mrg = 100)
    
  hexFoc = MyHexBin(focus, st_bbox(reference), cellsize = cellsize, mrg = 100) %>%
  mutate(expect = (nrow(focus) / nrow(reference)) * hexRef$ïAccUID) %>%
  mutate(chi = (ïAccUID-expect)/sqrt(expect),
         id = row_number()) %>%
  filter(!is.na(chi))
  
  quantDiff = quantile(hexFoc$chi, probs = seq(0,1,0.0025))
  
  if (mode == "view") {
    tm_shape(datapoints) +
      tm_dots(
        size = 0.01,
        alpha = pointopacity
      ) +
    tm_shape(hexFoc) +
      tm_polygons(
        col = "chi",
        id = "chi",
        palette = c("#053061", "#3885BC", "#A4CEE3", "#F7F7F7", "#F6B596", "#CB4A42", "#67001F"),
        alpha = 0.7,
        style = "fixed",
        breaks = c(min(hexFoc$chi), 
                   quantDiff[2], 
                   quantDiff[9], 
                   quantDiff[40], 
                   quantDiff[360], 
                   quantDiff[393], 
                   quantDiff[400], 
                   max(hexFoc$chi)),
        lwd = 0.1
      ) +
    tm_layout(title = title)
    
  } else if (mode == "return") {
      return(hexFoc)
    
  } else {
      warning("Mode has to be \"view\" or \"return\"!")
  }
}


```

Creating Hexbinning Count Maps of Bicycle, Motorcycle and Pedestrians
```{r}

BicCountMap <- countMap(MyHexBin, data = bycicle, title = "Bicycle Accidents Counts Map", datapoints = bycicle)

MotCountMap <- countMap(MyHexBin, data = motorCycle, title = "Motorcycle Accidents Counts Map", datapoints = motorCycle)

PedCountMap <- countMap(MyHexBin, data = pedestrians, title = "Pedestrians Accidents Counts Map", datapoints = pedestrians)

```

Visualize Hexbinning Count Maps of Bicycle, Motorcycle and Pedestrians
```{r}

BicCountMap
MotCountMap
PedCountMap

```


Creating Hexbinning Chi Maps of Bicycle, Motorcycle and Pedestrians
```{r}

BicChiMap <- chiMap(MyHexBin, reference = accidents, focus = bycicle, cellsize = 100, title = "Bicycle Accidents Chi Map", datapoints = bycicle, mode = "view") 

MotChiMap <- chiMap(MyHexBin, reference = accidents, focus = motorCycle, cellsize = 100, title = "Motorcycle Accidents Chi Map", datapoints = motorCycle, mode = "view") 

PedChiMap <- chiMap(MyHexBin, reference = accidents, focus = pedestrians, cellsize = 100, title = "Pedestrians Accidents Chi Map", datapoints = pedestrians, mode = "view") 

```

Visualizing Hexbinning Chi Maps of Bicycle, Motorcycle and Pedestrians
```{r}

BicChiMap
MotChiMap
PedChiMap

```

Creating Hexbinning Chi Maps of Severity-Categories
```{r}

SevereChiMap <- chiMap(MyHexBin, reference = accidents, focus = severe, cellsize = 100, title = "Severe Accidents Chi Map", datapoints = severe, mode = "view") 

```

Visualizing Hexbinning Chi Maps of Severity-Categories
```{r}

SevereChiMap

```

```{r, eval = FALSE}
diffMap = function(hexBinFunction, reference, focus, cellsize, title, mode = "view") {
  hexRef = hexBinFunction(reference, st_bbox(reference), cellsize = cellsize, mrg = 100) %>%
    mutate(countNorm = as.integer(ïAccUID/ nrow(reference) * 100000))
  
  hexFoc = hexBinFunction(focus, st_bbox(reference), cellsize = cellsize, mrg = 100) %>%
    mutate(countNorm = as.integer(ïAccUID/ nrow(focus) * 100000),
           expect = (nrow(focus) / nrow(reference)) * hexRef$ïAccUID)
  
  hexDiff = hexRef[,2] %>%
    mutate(countNorm = hexFoc$countNorm - hexRef$countNorm,
           counts = hexFoc$ïAccUID) %>%
    filter(countNorm != 0)
  
  quantDiff = quantile(hexDiff$countNorm, probs = seq(0,1,0.0025))
  
  if (mode == "view") {
    tm_shape(hexDiff) +
        tm_polygons(col = "countNorm",
                    title = "Difference to reference",
                    id = "counts",
                    palette = c("#053061", "#3885BC", "#A4CEE3", "#F7F7F7", "#F6B596", "#CB4A42", "#67001F"),
                    alpha = 0.7,
                    style = "fixed",
                    breaks = c(min(hexDiff$countNorm), 
                               quantDiff[2], 
                               quantDiff[9], 
                               quantDiff[40], 
                               quantDiff[360], 
                               quantDiff[393], 
                               quantDiff[400], 
                               max(hexDiff$countNorm)),
                    lwd = 0.1) +
      tm_layout(title = title)
    
  } else if (mode == "return") {
      return(hexDiff)
    
  } else {
      warning("Mode has to be \"view\" or \"return\"!")
  }
}

bicTmap = diffMap(hexBinFunction = MyHexBin, reference = accidents, focus = bycicle, cellsize = 100, title = "Bicycle", mode = "view")
motTmap = diffMap(hexBinFunction = MyHexBin, reference = accidents, focus = motorCycle, cellsize = 100, title = "Motorcycle", mode = "view")
pedTmap = diffMap(hexBinFunction = MyHexBin, reference = accidents, focus = pedestrians, cellsize = 100, title = "Pedestrians", mode = "view")

bicTmap
motTmap
pedTmap
```





